---
title: "train_users"
author: "Lisa Oshita"
date: "1/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Variables: 

* id: user id
* date_account_created: the date of account creation
* timestamp_first_active: timestamp of the first activity, note that it can be earlier than date_account_created or date_first_booking because a user can search before signing up
* date_first_booking: date of first booking
* gender
* age
* signup_method
* signup_flow: the page a user came to signup up from
* language: international language preference
* affiliate_channel: what kind of paid marketing
* affiliate_provider: where the marketing is e.g. google, craigslist, other
* first_affiliate_tracked: what was the first marketing the user interacted with before the signing up
* signup_app
* first_device_type
* first_browser
* country_destination: this is the target variable you are to predict

```{r}
# load the data 
# 213,451 users
dir <- file.path(getwd(),"data")
train_users <- read.csv(file.path(dir, "train_users_2.csv"))
# converting all factor vars to character
train_users <- dplyr::mutate_if(train_users, is.factor, as.character)
```

```{r}
# ===========================================================================
# date account created
# ===========================================================================
# separating year & month of date account created
train_users$dateCreated <- as.Date(train_users$date_account_created, format="%Y-%m-%d")
train_users$yearCreated <- format(train_users$dateCreated, "%Y")
train_users$monthCreated <- format(train_users$dateCreated, "%m")

ggplot(train_users, aes(x = yearCreated, fill = monthCreated)) + 
  geom_histogram(stat = "count") + 
  scale_y_continuous("number accounts created") + 
  scale_x_discrete("year account created") + 
  ggtitle("Number of accounts created per year")

# number of accounts created increases from 2010 - 2014
# accounts created up to June 2014

# weekday of account created
train_users$weekdayCreated <- weekdays(train_users$dateCreated)
# proportion of users who booked, grouped by weekday account was created on 
train_users %>% 
  group_by(weekdayCreated) %>%
  summarise(n = n(), propBooked = sum(book == 1)/n)
# day of the week an account was created doesn't seem to matter 
```

```{r}
# ===========================================================================
# timestamp first active 
# ===========================================================================
# first active year 
train_users$firstactive_y <- stringr::str_sub(train_users$timestamp_first_active, start = 1, end = 4)

# first active month 
train_users$firstactive_m <- stringr::str_sub(train_users$timestamp_first_active, start = 5, end = 6)

# first active day 
train_users$firstactive_d <- stringr::str_sub(train_users$timestamp_first_active, start = 7, end = 8)

# first active full date 
train_users$firstactive_date <- as.Date(paste(train_users$firstactive_y, "-", 
                                              train_users$firstactive_m, "-", 
                                              train_users$firstactive_d, sep = ""), 
                                        format = "%Y-%m-%d")

# difference between date first active and date first book 
train_users$active_booklag <- train_users$firstactive_date - train_users$firstBook
ggplot(train_users, aes(x = abs(active_booklag))) + 
  geom_histogram()
median(abs(train_users$active_booklag), na.rm = T)
mean(abs(train_users$active_booklag), na.rm = T)
max(abs(train_users$active_booklag), na.rm = T)
min(abs(train_users$active_booklag), na.rm = T)

# difference between date active and date account created 
train_users$acct_active_lag <- train_users$dateCreated - train_users$firstactive_date
ggplot(train_users, aes(x = abs(acct_active_lag))) + 
  geom_histogram()
median(abs(train_users$acct_active_lag), na.rm = T)
mean(abs(train_users$acct_active_lag), na.rm = T)
max(abs(train_users$acct_active_lag), na.rm = T)
min(abs(train_users$acct_active_lag), na.rm = T)
```


```{r}
# ===========================================================================
# date first booking
# ===========================================================================
# separating year & month of date of first book 
train_users$firstBook <- as.Date(train_users$date_first_booking, format="%Y-%m-%d")
train_users$firstBookYear <- format(train_users$firstBook, "%Y")
train_users$firstBookMonth <- format(train_users$firstBook, "%m")

# number of users who did not book 
sum(is.na(train_users$firstBook))/nrow(train_users)

# 58.3% of users did not book (NDF)

# ===========================================================================
# how long it took users to book (after creating account)
train_users$time_until_book <- train_users$firstBook - train_users$dateCreated

# distribution of time until first booking 
ggplot(train_users, aes(x = time_until_book)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(-10, 400))

median(train_users$time_until_book, na.rm = TRUE)
mean(train_users$time_until_book, na.rm = TRUE)
max(train_users$time_until_book, na.rm = TRUE)
min(train_users$time_until_book, na.rm = TRUE)
# median: 3 days 
# mean: 44.3625 days 
# max: 365 days 
# min: -349 days (why negative?)
# ===========================================================================
# time series plot of date first book 
firstbookDate <- train_users %>% select(date_first_booking) %>% filter(date_first_booking != "")
firstbookDF <- data.frame(date = names(table(firstbookDate$date_first_booking)),
                          count = unname(table(firstbookDate$date_first_booking)))
firstbookDF$date <- as.Date(firstbookDF$date, format="%Y-%m-%d")
ggplot(firstbookDF, aes(x = date, y = count.Freq)) + 
  geom_point() + 
  geom_smooth() + 
  scale_x_date(date_breaks = "6 months") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# seasonal pattern?
# spike in bookings between 3/2014 and 9/2014 (summer months?)
# spike appears in september for every year 
```

```{r}
# ===========================================================================
# age 
# ===========================================================================
# histogram of ages (up to 100)
ggplot(train_users, aes(x = age)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 100))
# ===========================================================================
# max age = 2014 (quite a few outliers)
# median age = 34
# mean age = 49.7

# age quantiles 
quantile(train_users$age, na.rm = TRUE)
# ===========================================================================
# boxplots of age for each country 
ggplot(train_users, aes(x = country_destination, y = age)) + 
  geom_boxplot() + 
  scale_y_continuous(limit = c(0, 60))

# median age range travelling to each country 
train_users %>% 
  group_by(country_destination) %>% 
  summarise(median_age = median(age, na.rm = TRUE))
# lowest median age for spain and portugal 

# ===========================================================================
# issues with the age variable 
# ===========================================================================
# 87990 missing ages (41.2%)

length(train_users$age[which(train_users$age > 110)])
# 801 users have ages > 110 
# 710 of those users set their age as 2014 
# some users put their birth year instead of age? (e.g 1995, 1936...)

# airbnb terms of use: must be 18 years or older to create an account/use the service 
# 158 users have ages < 18
train_users$age[which(train_users$age < 18)]
```

```{r}
# ===========================================================================
# other categorical vars 
# ===========================================================================
# indicator variable for if a user booked a trip
train_users$book <- 1
train_users$book[is.na(train_users$firstBook)] <- 0
# ===========================================================================
# function to plot bar graphs of categorical vars, color coded by whether or not the user booked
plot_cat <- function(var) {
  ggplot(train_users, aes(x = train_users[[var]], fill = as.factor(book))) + 
    geom_bar() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    ggtitle(var) + 
    scale_x_discrete(var) 
}
factorvars <- c("gender", "signup_method", "language", "affiliate_channel", "affiliate_provider",
                "first_affiliate_tracked", "signup_app", "first_device_type", "first_browser")
purrr::map(factorvars, ~plot_cat(.))
# uneven distribution for all categorical vars 
# ===========================================================================
# function to get the proportion of users in each level of categorical vars that booked a trip 
proportionBooked <- function(var) {
  train_users %>% 
    group_by(train_users[[var]]) %>%
    summarise(n = n(),proportionBooked = sum(book == 1)/n()) %>% 
    arrange(desc(proportionBooked))
}
purrr::map(factorvars, ~proportionBooked(.))
# signup method: only 18% of users who signed up with google booked a trip
#                much less than those who signed up with basic/FB
# affiliate channel: users of affiliate channel - content - book less than any other 
# affiliate provider: users of craigslist, direct, google book more than others
#               users of meetup, gsp, wayn book the least
# first affiliate tracked: users of marketing, untracked, linked book the most
#               users of NA, local ops book the least 
# signup app: Users of web book the most, users of android and iOS book the least 
# first device type: users of mac desktop/desktop/windows desktop book the most 
#               users of android phone book the least 
# ===========================================================================
# proportion travelling to each country  
train_users %>% 
  group_by(signup_method) %>% 
  summarise(n = n(), 
            propAU = sum(country_destination == "AU")/n,
            propCA = sum(country_destination == "CA")/n,
            propDE = sum(country_destination == "DE")/n,
            propES = sum(country_destination == "ES")/n,
            propFR = sum(country_destination == "FR")/n,
            propGB = sum(country_destination == "GB")/n,
            propIT = sum(country_destination == "IT")/n,
            propNDF = sum(country_destination == "NDF")/n,
            propNL = sum(country_destination == "NL")/n,
            propOther = sum(country_destination == "other")/n,
            propPT = sum(country_destination == "PT")/n,
            propUS = sum(country_destination == "US")/n
            )
```

```{r}
# ===========================================================================
# missing values 
# ===========================================================================
# function to count + find proportion of missing vals
count_NA <- function(var) {
  if (class(train_users[[var]]) == "numeric") {
      missing <- data.frame(var = var, 
                            total = sum(is.na(train_users[[var]])),
                            proportion = sum(is.na(train_users[[var]])) / nrow(train_users)
                            )
  } else {
    missing <- data.frame(var = var, 
                          total = sum(train_users[[var]] == ""),
                          proportion = sum(train_users[[var]] == "") / nrow(train_users)
                          )
  }
  return(missing)
}

# replacing unknowns with "" in gender var
train_users$gender <- stringr::str_replace(train_users$gender, 
                                           pattern = "-unknown-", 
                                           replacement = "")

purrr::map_df(c("gender", "age", "first_affiliate_tracked"), ~count_NA(.))

# ===========================================================================
# indicator variable for if age is missing 
train_users$provideAge <- 1
train_users$provideAge[is.na(train_users$age)] <- 0

train_users %>%
  group_by(ageincluded) %>%
  summarise(n = n(), propBook = sum(book == 1)/n)
# 23% of users who don't include their age book 
# 55% of users who do include their age book 
# ===========================================================================
# indicator variable for if user provided gender
train_users$provideGender <- 1
train_users$provideGender[train_users$gender == ""] <- 0

train_users %>% 
  group_by(provideGender) %>%
  summarise(n = n(), propBook = sum(book == 1)/n)
# 30% of users who did not provide their gender actually booked a trip 
# 50% of users who provided their gender actually booked a trip 
```

```{r}
# ===========================================================================
# country_destination 
# ===========================================================================
# AU: Australia, CA: Canada, DE: Germany, ES: Spain, FR: France, 
# GB: United Kingdom, IT: Italy, NL: Netherlands, PT: Portugal

# bar chart of countries 
ggplot(train_users, aes(x = country_destination)) + geom_bar()
# uneven distribution across countries 
# proportion of all users visiting each country 
train_users %>% 
  group_by(country_destination) %>% 
  summarise(n = n(), proportion = n/nrow(train_users))
# 58% - NDF 
# 30% - US

# bar chart of countries, excluding NDF (only for users that actually booked a trip)
ggplot(subset(train_users, book == 1), aes(x = country_destination)) + 
  geom_bar()
train_users %>% 
  filter(book == 1) %>% 
  group_by(country_destination) %>%
  summarise(n = n(), proportion = n/nrow(subset(train_users, book == 1)))
# 70% of users who did book, booked their trip to somewhere in the US
```



```{r}
# ===========================================================================
# boosting 
# ===========================================================================
library(gbm)

# subset data 
traindat <- subset(train_users, select = c(gender, age, signup_method, signup_flow, signup_app, 
                       affiliate_channel, affiliate_provider, first_affiliate_tracked, 
                       first_device_type, first_browser, country_destination))

# convert character vars to factors (gbm doesn't work with characters)
traindat <- dplyr::mutate_if(traindat, is.character, as.factor)

# fit model 
boostAirbnb <- gbm(country_destination ~ ., data = traindat, 
                   distribution = "multinomial", n.trees = 5, interaction.depth = 4)
# ===========================================================================
# predictions - returns probabilities
predictions <- predict(boostAirbnb, newdata = traindat, n.trees = 5, type = "response")

# assigns predictions to a class
train_users$predictions <- apply(predictions, 1, which.max)
levels <- data.frame(countries = levels(train_users$country_destination), num = 1:12)
train_users$predictions[train_users$predictions == 8] <- "NDF"
train_users$predictions[train_users$predictions == 12] <- "US"

# accuracy 
sum(train_users$predictions == train_users$country_destination) / nrow(train_users)
```




