---
title: "train_users"
author: "Lisa Oshita"
date: "1/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Variables: 

* id: user id
* date_account_created: the date of account creation
* timestamp_first_active: timestamp of the first activity, note that it can be earlier than date_account_created or date_first_booking because a user can search before signing up
* date_first_booking: date of first booking
* gender
* age
* signup_method
* signup_flow: the page a user came to signup up from
* language: international language preference
* affiliate_channel: what kind of paid marketing
* affiliate_provider: where the marketing is e.g. google, craigslist, other
* first_affiliate_tracked: what was the first marketing the user interacted with before the signing up
* signup_app
* first_device_type
* first_browser
* country_destination: this is the target variable you are to predict

```{r}
# load the data 
# 213,451 users
train_users <- read.csv(file.path(dir, "train_users_2.csv"))
# converting all factor vars to character
train_users <- dplyr::mutate_if(train_users, is.factor, as.character)
```

```{r}
# date account created
# separating year & month of date account created
train_users$dateCreated <- as.Date(train_users$date_account_created, format="%Y-%m-%d")
train_users$yearCreated <- format(train_users$dateCreated, "%Y")
train_users$monthCreated <- format(train_users$dateCreated, "%m")

ggplot(train_users, aes(x = yearCreated, fill = monthCreated)) + 
  geom_histogram(stat = "count") + 
  scale_y_continuous("number accounts created") + 
  scale_x_discrete("year account created") + 
  ggtitle("Number of accounts created per year")
# number of accounts created increases from 2010 - 2014
```

```{r}
# date first booking
# separating year & month of date of first book 
train_users$firstBook <- as.Date(train_users$date_first_booking, format="%Y-%m-%d")
train_users$firstBookYear <- format(train_users$firstBook, "%Y")
train_users$firstBookMonth <- format(train_users$firstBook, "%m")

# number of users who did not book 
sum(is.na(train_users$firstBook))/nrow(train_users)
# 58.3% of users did not book (NDF)
```

```{r}
# age 
ggplot(train_users, aes(x = age)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 100))
median(train_users$age, na.rm = TRUE)
# max age = 2014
# median age = 34
# mean age = 49.7
# 87990 missing ages (41.2%)

# proportion of missing ages for female, male, unknown and other
train_users %>% 
  group_by(gender) %>%
  summarise(n = n(), proportion_NAage = sum(is.na(age))/n)

# boxplots of age for each country 
ggplot(train_users, aes(x = country_destination, y = age)) + 
  geom_boxplot() + 
  scale_y_continuous(limit = c(0, 60))

# median age range travelling to each country 
train_users %>% 
  group_by(country_destination) %>% 
  summarise(median_age = median(age, na.rm = TRUE))
# lowest median age for spain and portugal 
```

```{r}
# function to plot bar graphs of categorical vars 
plot_cat <- function(var) {
  ggplot(train_users, aes(x = train_users[[var]])) + 
    geom_bar() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    ggtitle(var) + 
    scale_x_discrete(var) 
}
factorvars <- c("gender", "signup_method", "language", "affiliate_channel", "affiliate_provider",
                "first_affiliate_tracked", "signup_app", "first_device_type", "country_destination")
purrr::map(factorvars, ~plot_cat(.))

# proportion travelling to each country  
train_users %>% 
  group_by(signup_method) %>% 
  summarise(n = n(), 
            propAU = sum(country_destination == "AU")/n,
            propCA = sum(country_destination == "CA")/n,
            propDE = sum(country_destination == "DE")/n,
            propES = sum(country_destination == "ES")/n,
            propFR = sum(country_destination == "FR")/n,
            propGB = sum(country_destination == "GB")/n,
            propIT = sum(country_destination == "IT")/n,
            propNDF = sum(country_destination == "NDF")/n,
            propNL = sum(country_destination == "NL")/n,
            propOther = sum(country_destination == "other")/n,
            propPT = sum(country_destination == "PT")/n,
            propUS = sum(country_destination == "US")/n
            )
```

```{r}
# missing values 
# function to count + find proportion of missing vals
count_NA <- function(var) {
  if (class(train_users[[var]]) == "numeric") {
      missing <- data.frame(var = var, 
                            total = sum(is.na(train_users[[var]])),
                            proportion = sum(is.na(train_users[[var]])) / nrow(train_users)
                            )
  } else {
    missing <- data.frame(var = var, 
                          total = sum(train_users[[var]] == ""),
                          proportion = sum(train_users[[var]] == "") / nrow(train_users)
                          )
  }
  return(missing)
}

# replacing unknowns with "" in gender var
train_users$gender <- stringr::str_replace(train_users$gender, 
                                           pattern = "-unknown-", 
                                           replacement = "")

purrr::map_df(c("gender", "age", "first_affiliate_tracked"), ~count_NA(.))

# ===========================================================================

# indicator variable for if a user booked a trip
train_users$book <- 1
train_users$book[is.na(train_users$firstBook)] <- 0
# indicator variable for if age is missing 
train_users$provideAge <- 1
train_users$provideAge[is.na(train_users$age)] <- 0

train_users %>%
  group_by(ageincluded) %>%
  summarise(n = n(), propBook = sum(book == 1)/n)
# 23% of users who don't include their age book 
# 55% of users who do include their age book 

# indicator variable for if user provided gender
train_users$provideGender <- 1
train_users$provideGender[train_users$gender == ""] <- 0

train_users %>% 
  group_by(provideGender) %>%
  summarise(n = n(), propBook = sum(book == 1)/n)
# 30% of users who did not provide their gender actually booked a trip 
# 50% of users who provided their gender actually booked a trip 
```

```{r}
# country_destination 

# AU: Australia, CA: Canada, DE: Germany, ES: Spain, FR: France, 
# GB: United Kingdom, IT: Italy, NL: Netherlands, PT: Portugal

# bar chart of countries 
ggplot(train_users, aes(x = country_destination)) + geom_bar()
train_users %>% 
  group_by(country_destination) %>% 
  summarise(n = n(), proportion = n/nrow(train_users))
# 58% - NDF 
# 30% - US

# bar chart of countries, excluding NDF (only for users that actually booked a trip)
ggplot(subset(train_users, book == 1), aes(x = country_destination)) + 
  geom_bar()
train_users %>% 
  filter(book == 1) %>% 
  group_by(country_destination) %>%
  summarise(n = n(), proportion = n/nrow(subset(train_users, book == 1)))
# 70% of users who did book, booked their trip to somewhere in the US
```



```{r}
# ===========================================================================
# boosting 
# ===========================================================================
library(gbm)

# subset data 
traindat <- subset(train_users, select = c(gender, age, signup_method, signup_flow, signup_app, 
                       affiliate_channel, affiliate_provider, first_affiliate_tracked, 
                       first_device_type, first_browser, country_destination))

# convert character vars to factors (gbm doesn't work with characters)
traindat <- dplyr::mutate_if(traindat, is.character, as.factor)

# fit model 
boostAirbnb <- gbm(country_destination ~ ., data = traindat, 
                   distribution = "multinomial", n.trees = 5, interaction.depth = 4)
# ===========================================================================
# predictions - returns probabilities
predictions <- predict(boostAirbnb, newdata = traindat, n.trees = 5, type = "response")

# assigns predictions to a class
train_users$predictions <- apply(predictions, 1, which.max)
levels <- data.frame(countries = levels(train_users$country_destination), num = 1:12)
train_users$predictions[train_users$predictions == 8] <- "NDF"
train_users$predictions[train_users$predictions == 12] <- "US"

# accuracy 
sum(train_users$predictions == train_users$country_destination) / nrow(train_users)
```




